Verifying Hierarchical Cache Coherence Protocols: a theorem-prover based modular, reusable proof

### 2014-02-04 paper/System.tex
some modifications of background

The behavior of the system is described via the atomic state transitions. These
transitions have two parts: a) the guard dictating when the transition can
fire, and b) the action defining how the state of the system changes when the
transition happens. In the actual implementation, several of these transitions
or \emph{guarded atomic actions} happen concurrently in a single hardware
clock, but they will logically happen happen one at a time. Figure~\ref{trans}
gives the atomic transitions for the MIS protocol.
The actions in the transitions in Figure \ref{trans} show only the specific
states that change, the rest of the states remain as is. The notation ($x
\Leftarrow y$;) in a transition indicates setting of a specific state $x$ of
the overall system to value $y$.

@@

Figure~\ref{procside} shows the transitions that serve a processors request in
the leaf caches. LoadReq and StoreReq show the processing of a load and store
request, respectively. In the case of a store request, the data in the cache
for the requested address is updated with that supplied in the request. We do
not explicitly show the response generated by the cache in case of a load
request in our memory subsystem model; in a real system, the data in the cache
for the requested address is sent back to the processor. Our definition of
store atomicity requires responses for both load and store requests, and the
response field contains the $(\textit{Processor}*\mathbb{N})$ pair identifying
the request corresponding to the response, the data from the cache in case of a
load response, and a value for the field \tme. A \Response{} is defined
whenever LoadReq or StoreReq transition takes place. Let $c$ be a leaf cache,
and $t$ be the position of a LoadReq$(c)$ or StoreReq$(c)$ transition in the
ordered list of transitions of the full system starting from the initial state.
The fields of the response are as follows:
\begin{multline*}
\small\langle (c, c.\pos), t, \mylet q := \reqFn(c, c.\pos) \myin\\
\small\myif q.\desc = \Ld \mythen c.\data[q.\addr] \myelse \_ \rangle
\end{multline*}

@@

Note that the atomic transitions in Figure~\ref{trans} access only local states,
\ie{} a transition can read or write states only corresponding to a single cache
and/or the channel the cache is connected to. This restriction allows such a
system to be implemented directly into hardware. Bluespec System Verilog (BSV),
for instance, directly converts these transitions into efficient synchronous
hardware. Though these transitions logically happen one by one, BSV executes
several transitions simultaneously, as long as they do not access the same
state, \ie{} do not \emph{conflict}~\cite{Hoe:TCAD,HoeArvind:TRSSynthesis1}.
Sometimes, these transitions (both the guards and the actions) may not be
amenable to single cycle implementations in hardware. For example, writing a
register array (like a cache) can actually take several hardware clock cycles,
though logically it is a single transition.  Karczmarek \etal~\cite{Karczmarek}
has provided a scheme to convert atomic transitions into synchronous hardware
in which each transition can span several clock cycles.

>>>>

The behavior of the system is described via the atomic state transitions. These
transitions have two parts: a) the guard dictating when the transition can
fire, and b) the action defining how the state of the system changes when the
transition happens. Logically, these transitions happen one at a time.
Figure~\ref{trans} gives the atomic transitions for the MIS protocol.  The
actions in the transitions in Figure \ref{trans} show only the specific states
that change, the rest of the states remain as is. The notation ($x \Leftarrow
y$;) in a transition indicates setting of a specific state $x$ of the overall
system to value $y$.

@@

Figure~\ref{procside} shows the transitions that serve a processors request in
the leaf caches. LoadReq and StoreReq show the processing of a load and store
request, respectively. In the case of a store request, the data in the cache
for the requested address is updated with that supplied in the request. We do
not explicitly show the response generated by the cache in case of a load
request; in a real system, the data in the cache for the requested address is
sent back to the processor. When a LoadReq or StoreReq happens at a cache $c$
pertaining to address $a$, the \Response{}, needed in the definition of store
atomicity, is generated as follows: the \labelR{} field contains $(c, c.\pos)$
pair, the \dataR{} field contains the value $c.\data[a]$, and \timeR{} field
contains the number of transitions that has taken place in the system before
the current transition, intuitively denoting the time at which this transition
takes place.

@@

Note that the atomic transitions in Figure~\ref{trans} access only local states,
\ie{} a transition can read or write states only corresponding to a single cache
and/or the channel the cache is connected to. This restriction allows such a
system to be implemented directly into hardware. Bluespec System Verilog (BSV),
for instance, directly converts these transitions into efficient synchronous
hardware. Though these transitions logically happen one by one, BSV executes
several transitions simultaneously, as long as they do not access the same
state, \ie{} do not \emph{conflict}~\cite{Hoe:TCAD,HoeArvind:TRSSynthesis1}.
Sometimes, these transitions (both the guards and the actions) may not be
amenable to single cycle implementations in hardware. For example, writing a
register array can actually take several hardware clock cycles, though
logically it is a single transition.  Karczmarek \etal~\cite{Karczmarek} has
provided a scheme to convert atomic transitions into synchronous hardware in
which each transition can span several clock cycles.
